---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  results = 'hold',
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![](https://cranlogs.r-pkg.org/badges/yfR)](https://CRAN.R-project.org/yfR)

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

# Motivation

`yfR` is the second and backwards-incompatible version of [BatchGetSymbols](https://CRAN.R-project.org/package=BatchGetSymbols). It provides access to daily stock prices from [Yahoo Finance](https://finance.yahoo.com/), a vast repository with data around the globe. Moreover, *yfR* allows large scales downloads of data, using a local caching system for speeding up the process.

## Features

- Fetchs daily/weekly/monthly/annual stock prices (and returns) from yahoo finance and returns a dataframe in the long format;
- A session-persistent smart cache system is available by default. This means that the data is saved locally and only missing portions are downloaded, if needed.
- All dates are compared to a benchmark ticker such as SP500 and, whenever an individual asset does not have a sufficient number of dates, the software drops it from the output. This means you can choose to ignore tickers with high number of missing dates.
- A customized function called `yf_convert_to_wide()` can transform the long dataframe into a wide format (tickers as columns). 
- Parallel computing is available, speeding up the data importation process.


## Warnings

- Yahoo finance data is far from perfect or reliable, specially for individual stocks. In my experience, using it for research code with stock **indices** is fine and I can match it with other data sources. But, adjusted stock prices for **individual assets** is messy as stock events such as splits or dividends are not properly registered. I was never able to match it with other data sources, specially for long time periods with lots of corporate events. My advice is to never use the data of individual stocks in production (research papers or academic documents -- thesis and dissertations). 


## Installation

```
# CRAN (not yet available)
#install.packages('yfR')

# Github (dev version)
devtools::install_github('msperlin/yfR')
```


## Examples

### Fetching a single stock price

```{r, results='hold'}
library(yfR)

my_ticker <- 'FB'
first_date <- Sys.Date() - 30
last_date <- Sys.Date()

df_yf <- yf_get_data(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

str(df_yf)
```

### Fetching many stock prices

```{r}
library(yfR)
library(ggplot2)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get_data(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

str(df_yf_multiple)

p <- ggplot(df_yf_multiple, aes(x = ref_date, y = price_adjusted,
                       color = ticker)) + 
  geom_line()

p

```


### Fetching collections of prices

Collections are just a bundle of tickers pre-organized in the package. For example, collection `SP500` represents the current composition of the SP500 index. 

```{r, eval=FALSE}
library(yfR)

df_yf <- yf_get_collection("SP500")

str(df_yf)
```


### Fetching daily/weekly/monthly/yearly price data 

```{r}
library(yfR)
library(ggplot2)
library(dplyr)

my_ticker <- 'GE'
first_date <- '2010-01-01'
last_date <- Sys.Date()

df_dailly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'daily') |>
  mutate(freq = 'daily')
  
  
df_weekly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'weekly') |>
  mutate(freq = 'weekly')

df_monthly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'monthly') |>
  mutate(freq = 'monthly')

df_yearly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'yearly') |>
  mutate(freq = 'yearly')

df_allfreq <- bind_rows(
  list(df_dailly, df_weekly, df_monthly, df_yearly)
) |>
  mutate(freq = factor(freq, 
                       levels = c('daily', 
                                  'weekly',
                                  'monthly',
                                  'yearly'))) # make sure the order in plot is right

p <- ggplot(df_allfreq, aes(x=ref_date, y = price_adjusted)) + 
  geom_point() + geom_line() + facet_grid(freq ~ ticker) + 
  theme_minimal() + 
  labs(x = '', y = 'Adjusted Prices')

print(p)
```


### Changing format to wide

```{r}
library(yfR)
library(ggplot2)
library(kableExtra)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get_data(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

l_wide <- yf_converto_to_wide(df_yf_multiple)

prices_wide <- l_wide$price_adjusted

knitr::kable(head(prices_wide)) 
```

