---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  results = 'hold',
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![](https://cranlogs.r-pkg.org/badges/yfR)](https://CRAN.R-project.org/yfR)

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

# Motivation

`yfR` is the second and backwards-incompatible version of [BatchGetSymbols](https://CRAN.R-project.org/package=BatchGetSymbols). In a nutshell, it provides access to daily stock prices from [Yahoo Finance](https://finance.yahoo.com/), a vast repository with financial data around the globe. Moreover, *yfR* allows large scale download of data using a local caching system and parallel computing for speeding up the process.


## Features

- Fetchs daily/weekly/monthly/annual stock prices (and returns) from yahoo finance and returns a dataframe in the long format (stacked data);

- A new feature called "collections" facilitates download of multiple tickers from a particular market/index. You can, for example, download data for all stocks in the SP500 index with a simple call to `yf_get_collection()`;

- A session-persistent smart cache system is available by default. This means that the data is saved locally and only missing portions are downloaded, if needed.

- All dates are compared to a benchmark ticker such as SP500 and, whenever an individual asset does not have a sufficient number of dates, the software drops it from the output. This means you can choose to ignore tickers with high number of missing dates.

- A customized function called `yf_convert_to_wide()` can transform the long dataframe into a wide format (tickers as columns). The output is a list where each element is a different target variable (prices, returns, volumes).

- Parallel computing is available, speeding up the data importation process.

## Differences from [BatchGetSymbols](https://github.com/msperlin/BatchGetSymbols)

Package `BatchgetSymbols` was developed in 2016, with bad choices from my part. As I learned more about R, my code became better and easier to maintain. However, it is impossible to keep compatibility with the changes I wanted to make, which is why I created a new package. 

Here are the main differences from one to the other: 

- All input arguments are now formatted as "snake_case" and not "dot.case". For example, the argument for the first date of data importation in `yfR::get_yf_data` is `first_date`, and not `first.date` (see arguments in `BatchGetSymbols::BatchGetSymbols`)

- All function have been renamed for a common API notation. For example, `BatchGetSymbols::BatchGetSymbols` is now `yfR::get_yf_data`.

- The output is always a tibble with the price data (and not a list). If one wants the tibble with a summary of the importing process, it is available as an attribute of the output (see function `base::attributes`)

- A new feature called "collection", which allows for easy download of a collection of tickers. For example, you can download price data for all components of the SP500 by simply calling `yfR::yf_get_collection("SP500")`.

- New and prettier status messages using package `cli`


## Warnings

- Yahoo finance data is far from perfect or reliable, specially for individual stocks. In my experience, using it for research code with stock **indices** is fine and I can match it with other data sources. But, adjusted stock prices for **individual assets** is messy as stock events such as splits or dividends are not properly registered. I was never able to match it with other data sources, specially for long time periods with lots of corporate events. My advice is to **never use the yahoo finance data of individual stocks in production** (research papers or academic documents -- thesis and dissertations). 


## Installation

```
# CRAN (not yet available)
#install.packages('yfR')

# Github (dev version)
devtools::install_github('msperlin/yfR')
```


## Examples

### Fetching a single stock price

```{r, results='hold'}
library(yfR)

# set options for algorithm
my_ticker <- 'FB'
first_date <- Sys.Date() - 30
last_date <- Sys.Date()

# fetch data
df_yf <- yf_get_data(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

# output is a tibble with data
str(df_yf)
```

### Fetching many stock prices

```{r}
library(yfR)
library(ggplot2)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get_data(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

str(df_yf_multiple)

p <- ggplot(df_yf_multiple, aes(x = ref_date, y = price_adjusted,
                       color = ticker)) + 
  geom_line()

p

```


### Fetching collections of prices

Collections are just a bundle of tickers pre-organized in the package. For example, collection `SP500` represents the current composition of the SP500 index. 

```{r, eval=FALSE}
library(yfR)

df_yf <- yf_get_collection("SP500")

str(df_yf)
```


### Fetching daily/weekly/monthly/yearly price data 

```{r}
library(yfR)
library(ggplot2)
library(dplyr)

my_ticker <- 'GE'
first_date <- '2010-01-01'
last_date <- Sys.Date()

df_dailly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'daily') |>
  mutate(freq = 'daily')
  
  
df_weekly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'weekly') |>
  mutate(freq = 'weekly')

df_monthly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'monthly') |>
  mutate(freq = 'monthly')

df_yearly <- yf_get_data(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'yearly') |>
  mutate(freq = 'yearly')

df_allfreq <- bind_rows(
  list(df_dailly, df_weekly, df_monthly, df_yearly)
) |>
  mutate(freq = factor(freq, 
                       levels = c('daily', 
                                  'weekly',
                                  'monthly',
                                  'yearly'))) # make sure the order in plot is right

p <- ggplot(df_allfreq, aes(x=ref_date, y = price_adjusted)) + 
  geom_point() + geom_line() + facet_grid(freq ~ ticker) + 
  theme_minimal() + 
  labs(x = '', y = 'Adjusted Prices')

print(p)
```


### Changing format to wide

```{r}
library(yfR)
library(ggplot2)
library(kableExtra)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get_data(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

l_wide <- yf_converto_to_wide(df_yf_multiple)

prices_wide <- l_wide$price_adjusted

knitr::kable(head(prices_wide)) 
```

