---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  results = 'hold',
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Codecov test coverage](https://codecov.io/gh/msperlin/yfR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/msperlin/yfR?branch=main)
[![R build (rcmdcheck)](https://github.com/msperlin/yfR/workflows/R-CMD-check/badge.svg)](https://github.com/msperlin/yfR/actions)

<!-- badges: end -->

# Motivation

`yfR` is the second and backwards-incompatible version of [BatchGetSymbols](https://CRAN.R-project.org/package=BatchGetSymbols). In a nutshell, it provides access to daily stock prices from [Yahoo Finance](https://finance.yahoo.com/), a vast repository with financial data around the globe. Yahoo Finance cover a large number of markets and assets, being used extensively for importing price datasets used in academic research and teaching. 

Package `yfR` is based on [quantmod](https://www.quantmod.com/) and used its main function for fetching data from Yahoo Finance. The main innovation in `yfR` is in the organization of the imported financial data and using local caching system and parallel computing for speeding up large scale download of datasets from Yahoo Finance.

See full documentation [here](https://github.com/msperlin/yfR).

## Features

- Fetchs daily/weekly/monthly/annual stock prices/returns from yahoo finance and outputs a dataframe (tibble) in the long format (stacked data);

- A new feature called "collections" facilitates download of multiple tickers from a particular market/index. You can, for example, download data for all stocks in the SP500 index with a simple call to `yf_collection_get()`;

- A session-persistent smart cache system is available by default. This means that the data is saved locally and only missing portions are downloaded, if needed.

- All dates are compared to a benchmark ticker such as SP500 and, whenever an individual asset does not have a sufficient number of dates, the software drops it from the output. This means you can choose to ignore tickers with high number of missing dates.

- A customized function called `yf_convert_to_wide()` can transform the long dataframe into a wide format (tickers as columns), much used in portfolio optimization. The output is a list where each element is a different target variable (prices, returns, volumes).

- Parallel computing with package `furrr` is available, speeding up the data importation process.


## Differences from [BatchGetSymbols](https://github.com/msperlin/BatchGetSymbols)

Package `BatchgetSymbols` was developed back in 2016, with many bad structural choices from my part. Since then, I learned more about R and its ecosystem, resulting in better and more maintainable code. However, it is impossible to keep compatibility with the changes I wanted to make, which is why I decided to develop a new (and fresh) package. 

Here are the main differences between `yfR` (new) and `BatchGetSymbols` (old): 

- All input arguments are now formatted as "snake_case" and not "dot.case". For example, the argument for the first date of data importation in `yfR::yf_get()` is `first_date`, and not `first.date` as used in  `BatchGetSymbols::BatchGetSymbols`.

- All function have been renamed for a common API notation. For example, `BatchGetSymbols::BatchGetSymbols` is now `yfR::yf_get()`. Likewise, the function for fetching collections is `yfR::yf_collection_get()`.

- The output of `yfR::yf_get()` is always a tibble with the price data (and not a list as in `BatchGetSymbols::BatchGetSymbols`). If one wants the tibble with a summary of the importing process, it is available as an attribute of the output (see function `base::attributes`)

- A new feature called "collection", which allows for easy download of a collection of tickers. For example, you can download price data for all components of the SP500 by simply calling `yfR::yf_collection_get("SP500")`.

- New and prettier status messages using package `cli`


## Warnings

- Yahoo finance data is far from perfect or reliable, specially for individual stocks. In my experience, using it for research code with stock **indices** is fine and I can match it with other data sources. But, adjusted stock prices for **individual assets** is messy as stock events such as splits or dividends are not properly registered. I was never able to match it with other data sources, specially for long time periods with lots of corporate events. My advice is to **never use the yahoo finance data of individual stocks in production** (research papers or academic documents -- thesis and dissertations). If adjusted price data of individual stocks is important for your research, **use other data sources** such as [EOD](https://eodhistoricaldata.com/), [SimFin](https://simfin.com/) or [Economática](https://economatica.com/).


## Installation

```
# CRAN (not yet available)
#install.packages('yfR')

# Github (dev version)
devtools::install_github('msperlin/yfR')
```


## Examples

### Fetching a single stock price

```{r, results='hold'}
library(yfR)

# set options for algorithm
my_ticker <- 'FB'
first_date <- Sys.Date() - 30
last_date <- Sys.Date()

# fetch data
df_yf <- yf_get(tickers = my_ticker, 
                     first_date = first_date,
                     last_date = last_date)

# output is a tibble with data
head(df_yf)
```


### Fetching many stock prices

```{r}
library(yfR)
library(ggplot2)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get(tickers = my_ticker, 
                              first_date = first_date,
                              last_date = last_date)


p <- ggplot(df_yf_multiple, 
            aes(x = ref_date, y = price_adjusted,
                color = ticker)) + 
  geom_line()

print(p)

```


### Fetching collections of prices

Collections are just a bundle of tickers pre-organized in the package. For example, collection `SP500` represents the current composition of the SP500 index. 

```{r, eval=FALSE}
library(yfR)

df_yf <- yf_collection_get("SP500", 
                           first_date = Sys.Date() - 30,
                           last_date = Sys.Date())

head(df_yf)
```


### Fetching daily/weekly/monthly/yearly price data 

```{r}
library(yfR)
library(ggplot2)
library(dplyr)

my_ticker <- 'GE'
first_date <- '2010-01-01'
last_date <- Sys.Date()

df_dailly <- yf_get(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'daily') %>%
  mutate(freq = 'daily')


df_weekly <- yf_get(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'weekly') %>%
  mutate(freq = 'weekly')

df_monthly <- yf_get(tickers = my_ticker, 
                          first_date, last_date, 
                          freq_data = 'monthly') %>%
  mutate(freq = 'monthly')

df_yearly <- yf_get(tickers = my_ticker, 
                         first_date, last_date, 
                         freq_data = 'yearly') %>%
  mutate(freq = 'yearly')

df_allfreq <- bind_rows(
  list(df_dailly, df_weekly, df_monthly, df_yearly)
) %>%
  mutate(freq = factor(freq, 
                       levels = c('daily', 
                                  'weekly',
                                  'monthly',
                                  'yearly'))) # make sure the order in plot is right

p <- ggplot(df_allfreq, aes(x=ref_date, y = price_adjusted)) + 
  geom_point() + geom_line() + facet_grid(freq ~ ticker) + 
  theme_minimal() + 
  labs(x = '', y = 'Adjusted Prices')

print(p)
```


### Changing format to wide

```{r}
library(yfR)
library(ggplot2)
library(kableExtra)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get(tickers = my_ticker, 
                              first_date = first_date,
                              last_date = last_date)

l_wide <- yf_convert_to_wide(df_yf_multiple)

prices_wide <- l_wide$price_adjusted

head(prices_wide)
```

