---
title: "Getting Started"
author: "Marcelo Perlin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Examples

Here you'll find a series of example of calls to `yf_get()`. Most arguments
are self-explanatory, but you can find more details at the help files.

The steps of the algorithm are:

1. check cache files for existing data
2. if not in cache, fetch stock prices from YF and clean up the raw data
3. write cache file if not available
4. calculate all returns
5. build diagnostics
6. return the data to the user

### Fetching a single stock price

```{r, results='hold'}
library(yfR)

# set options for algorithm
my_ticker <- 'FB'
first_date <- Sys.Date() - 30
last_date <- Sys.Date()

# fetch data
df_yf <- yf_get(tickers = my_ticker, 
                first_date = first_date,
                last_date = last_date)

# output is a tibble with data
head(df_yf)
```


### Fetching many stock prices

```{r}
library(yfR)
library(ggplot2)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get(tickers = my_ticker, 
                         first_date = first_date,
                         last_date = last_date)


p <- ggplot(df_yf_multiple, aes(x = ref_date, y = price_adjusted,
                                color = ticker)) + 
  geom_line()

p

```


### Fetching daily/weekly/monthly/yearly price data 

```{r}
library(yfR)
library(ggplot2)
library(dplyr)

my_ticker <- 'GE'
first_date <- '2010-01-01'
last_date <- Sys.Date()

df_dailly <- yf_get(tickers = my_ticker, 
                    first_date, last_date, 
                    freq_data = 'daily') %>%
  mutate(freq = 'daily')


df_weekly <- yf_get(tickers = my_ticker, 
                    first_date, last_date, 
                    freq_data = 'weekly') %>%
  mutate(freq = 'weekly')

df_monthly <- yf_get(tickers = my_ticker, 
                     first_date, last_date, 
                     freq_data = 'monthly') %>%
  mutate(freq = 'monthly')

df_yearly <- yf_get(tickers = my_ticker, 
                    first_date, last_date, 
                    freq_data = 'yearly') %>%
  mutate(freq = 'yearly')

df_allfreq <- bind_rows(
  list(df_dailly, df_weekly, df_monthly, df_yearly)
) %>%
  mutate(freq = factor(freq, 
                       levels = c('daily', 
                                  'weekly',
                                  'monthly',
                                  'yearly'))) # make sure the order in plot is right

p <- ggplot(df_allfreq, aes(x=ref_date, y = price_adjusted)) + 
  geom_point() + 
  geom_line() + 
  facet_grid(freq ~ ticker) + 
  theme_minimal() + 
  labs(x = '', y = 'Adjusted Prices')

print(p)
```


### Changing format to wide

```{r}
library(yfR)
library(ggplot2)

my_ticker <- c('FB', 'GM', 'MMM')
first_date <- Sys.Date() - 100
last_date <- Sys.Date()

df_yf_multiple <- yf_get(tickers = my_ticker, 
                         first_date = first_date,
                         last_date = last_date)

print(df_yf_multiple)

l_wide <- yf_convert_to_wide(df_yf_multiple)

names(l_wide)

prices_wide <- l_wide$price_adjusted
head(prices_wide)
```
